---
output: 
  github_document
  # html_document:
    # self_contained: no
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# R package bfsl: Best-fit Straight Line
<!-- badges: start -->
[![R build status](https://github.com/pasturm/bfsl/workflows/R-CMD-check/badge.svg)](https://github.com/pasturm/bfsl/actions)
[![codecov](https://codecov.io/gh/pasturm/bfsl/branch/master/graph/badge.svg)](https://codecov.io/gh/pasturm/bfsl)
[![CRAN version](https://www.r-pkg.org/badges/version-last-release/bfsl)](https://cran.r-project.org/package=bfsl)
[![CRAN downloads/month](https://cranlogs.r-pkg.org/badges/bfsl)](https://cran.r-project.org/package=bfsl)
[![CRAN total downloads](https://cranlogs.r-pkg.org/badges/grand-total/bfsl)](https://cran.r-project.org/package=bfsl)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

### How to fit a straight line through a set of points with errors in both coordinates?

The solution for the best-fit straight line to independent points with normally distributed errors in both _x_ and _y_ is known e.g. from York ([1966](#york66), [1968](#york68), [2004](#york04)). It provides unbiased estimates of the intercept, slope and standard errors of the best-fit straight line, even when the _x_ and _y_ errors are correlated.

The bfsl package implements York's general solution and provides the best-fit straight line of bivariate data with errors in both coordinates.

Other commonly used least-squares estimation methods, such as ordinary least-squares regression, orthogonal distance regression (also called major axis regression), geometric mean regression (also called reduced major axis or standardised major axis regression) or Deming regression are all special cases of York’s solution and only valid under particular measurement conditions.


## Example
<!-- # ```{r print} -->
<!-- # library(bfsl) -->
<!-- # bfsl(pearson_york_data) -->
<!-- # ``` -->
```{r plot, out.width="75%", dpi=300}
library(bfsl)
fit = bfsl(pearson_york_data)
print(fit)

plot(fit)
ols = bfsl(pearson_york_data, sd_x = 0, sd_y = 1)
abline(coef = ols$coef[,1], lty = 2)
legend("topright", c("ordinary least squares", "best-fit straight line"), lty = c(2,1))
```

Confindence interval:
```{r confidence, out.width="75%", dpi=300}
df = as.data.frame(fit$data)
newx = seq(min(df$x-df$sd_x), max(df$x+df$sd_x), length.out = 100)
preds = predict(fit, newdata = data.frame(x=newx), interval = 'confidence')

# plot
plot(y ~ x, data = df, type = 'n', xlim = c(min(df$x-df$sd_x), max(df$x+df$sd_x)),
     ylim = c(min(df$y-df$sd_y), max(df$y+df$sd_y)), las = 1)
grid()
polygon(c(rev(newx), newx), c(rev(preds[ ,3]), preds[ ,2]), col = 'grey90', border = NA)
abline(coef = fit$coef[,1], lty = 1)
points(df$x, df$y)
arrows(df$x, df$y-df$sd_y, df$x, df$y+df$sd_y, length = 0.05, angle = 90, code = 3)
arrows(df$x-df$sd_x, df$y, df$x+df$sd_x, df$y, length = 0.05, angle = 90, code = 3)
```
```{r ggplot, out.width="60%", dpi=300}
# with ggplot2
library(ggplot2)
ggplot(data = df, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = bfsl) +
  geom_errorbar(aes(ymin = y-sd_y, ymax = y+sd_y), width = 0.05) +
  geom_errorbarh(aes(xmin = x-sd_x, xmax = x+sd_x), height = 0.05)
```

## Installation
Install bfsl from CRAN with:
```{r cran_install, eval=FALSE}
install.packages("bfsl")
```
Install the development version from GitHub with:
```{r github_install, eval=FALSE}
if (!require("remotes")) { install.packages("remotes") }
remotes::install_github("pasturm/bfsl")
```

See the [NEWS file](https://github.com/pasturm/bfsl/blob/master/NEWS.md) for latest release notes.

## References

<a name="york66"></a>York, D. (1966). Least-squares fitting of a straight line. _Canadian Journal of Physics_, 44(5), 1079–1086, https://doi.org/10.1139/p66-090

<a name="york68"></a>York, D. (1968). Least squares fitting of a straight line with correlated errors. _Earth and Planetary Science Letters_, 5, 320–324,
https://doi.org/10.1016/S0012-821X(68)80059-7

<a name="williamson68"></a>Williamson, J. H. (1968). Least-squares fitting of a straight line,
_Canadian Journal of Physics_, 46(16), 1845-1847, https://doi.org/10.1139/p68-523

<a name="york04"></a>York, D. et al. (2004). Unified equations for the slope, intercept, and 
standard errors of the best straight line, _American Journal of Physics_, 72, 367-375, https://doi.org/10.1119/1.1632486

<a name="cantrell08"></a>Cantrell, C. A. (2008). Technical Note: Review of methods 
for linear least-squares fitting of data and application to atmospheric chemistry problems,
_Atmospheric Chemistry and Physics_, 8, 5477-5487, https://www.atmos-chem-phys.net/8/5477/2008/

<a name="wehr17"></a>Wehr, R. and Saleska, S. R. (2017). The long-solved problem of the best-fit straight line: application to isotopic mixing lines, _Biogeosciences_, 14, 17-29, https://doi.org/10.5194/bg-14-17-2017
